The relativistic capture of compact objects by intermediate-mass black holes

### 2019-06-03 mn.tex
abstract and introduction revised

Intermediate mass black holes, with masses in the range $10^2-10^5\Ms$, might
represent the missing link between stellar and supermassive mass black holes.
Dense stellar systems, such as globular clusters (GCs), are thought to be ideal
factories for IMBH production via formation and collapse of a very massive
stars through stellar collisions \citep{zwart02, giersz15, mapelli16} or via
multiple interactions and mergers between stars and stellar-mass BHs
\citep{giersz15}. Unfortunately, a striking observational evidence for the
presence of IMBHs inhabiting GCs is still missing due to the the little
dynamical effects that these objects have on their surroundings. Indeed, models
suggest that several processes can mimic an IMBH, like anisotropies in the GC
kinematical properties \citep{zocchi}, or the presence of a dense cluster of
stellar mass BHs that dominate the inner cluster centre
\citep{AAG18a,AAG18b,AS16,vandermarel10}. Nevertheless, a few observational
IMBHs candidates have been found for Galactic GCs
\citep{noyola10,lu13,lanzoni13,kiziltan17} and their mass and influence radius
might be possibly connected with the host cluster observational properties
\citep{AAG18a}. Due to this, finding an unique way to unravel the presence of
IMBHs in GCs represents one of the most interesting challenges in modern
astronomy. Beside this, the presence of an IMBH sitting in the centre of a
dense cluster represents a scenario particularly appealing from the
perspectives of gravitational waves (GW) astronomy. Indeed, a compact remnants
orbiting the IMBH can enter a regime where GWs emission dominates, leading to
the formation of an intermediate mass-ratio inspiral
\citep[IMRI,][]{konstantinidis13,haster16,leigh14}, a class of sources possible
audible with the next generation of GW observatories like LISA
\citep{seoane07,amaro12,seoane18}

However, in the highly dense regions that characterize star clusters centres,
the formation of IMRIs is not a smooth process. Indeed, due to the continuous
interactions with stars, an IMRI ``progenitor'', namely a tight IMBH-BH binary,
might be subjected to strong perturbation induced, for instance by a passing-by
BH. The three-body interaction involving the IMBH and the two BHs can lead to a
variety of end states, including the formation of an IMRI, a stellar BH binary,
the ejection of one BH, or even both, or the development of a head-on
collision.

Quantifying the branching ratio for IMRIs formation constitute a fundamental
step to assess the probability to observe these GW sources with the next
generation of space-based detectors like
LISA\footnote{\url{https://www.elisascience.org/}} \citep{seoane07}, TianQin
\citep{tianqin16} or Taiji \citep{taiji17}.

In this paper, we model the potential formation of an IMRI from the
interactions between an IMBH and two stellar mass BHs. To reach the aim, we use
$N$-body simulations that take into account in particles' equations of motion
both the star cluster gravitational potential and post-Newtonian corrections up
to 2.5 order. Varying the IMBH and BHs masses, their orbital configuration, and
the host cluster structural properties, we build-up three sets consisting of
2000 models each, which allow us to uncover three possible scenarios for IMRIs
formation.

The paper is organized as follows: in Section \ref{num} we present and
summarize the numerical setup used to model the IMBH - BH interaction, in
Section \ref{met} we discuss the main properties of the simulations outputs,
and the implications for GW astronomy, while Section \ref{res} is devoted to
the results summary and conclusions.

>>>>

Intermediate-mass black holes, with masses in the range $10^2-10^5\Ms$
represent the missing link between stellar-mass and supermassive black holes.
Dense stellar systems, such as globular clusters (GCs), could be the place
where IMBH form via the creation of a very massive stars through a runaway
process which would eventually collapse
\citep{PortegiesZwartMcMillan00,GurkanEtAl04,PortegiesZwartEtAl04,
FreitagEtAl06, giersz15, mapelli16}.  This process however depends on a number
of assumptions, such as the metallicity, stellar winds, and the actualy outcome
of the stellar collisions
\citep[see][]{BelkusEtAl07,SuzukiEtAl07,GlebbeekEtAl09}.
Another possibility is through
multiple interactions and mergers between stars and stellar-mass BHs
\citep{giersz15}. Unfortunately, a striking observational evidence for the
presence of IMBHs inhabiting GCs is still missing due to the little
dynamical effects that these objects have on their surroundings. Indeed, models
suggest that several processes can mimic the presence of an IMBH, like anisotropies in the
kinematical properties of the GC \citep{zocchi}, or the presence of a dense cluster of
stellar-mass black holes that dominate the inner cluster centre
\citep{AAG18a,AAG18b,AS16,vandermarel10}. Nevertheless, a few observational
IMBHs candidates have been found for Galactic GCs
\citep{noyola10,lu13,lanzoni13,kiziltan17} and their mass and influence radius
might be possibly connected with the host cluster observational properties
\citep{AAG18a}.

It has been shown in the literature that a compact object, such as a
stellar-mass black hole, could be gravitationally captured by the IMBH, to
eventually merge with it. The first numerical simulation of this scenario was
presented by \cite{konstantinidis13}, who used dedicated direct-summation
$N-$body simulations with all relativistic phenomena included at least at the
lowest order (periapsis shift, energy loss due to gravitational radiation, and
the relativistic recoil). They observed an intermediate-mass ratio inspiral
\citep{Amaro-SeoaneLRR2012,Amaro-SeoaneEtAl07} forming and follow it down to
the merger, which leads to the ejection of the IMBH. Other studies have
followed their findings, in particular the work of
\cite{leigh14,MacLeodEtAl2016}; also, because the initial setup of
\cite{haster16} is almost identical to that of \cite{konstantinidis13} but they
use a different integrator, it can be envisaged as a confirmation of their
results. This source of gravitational waves (GWs) can be observed with a the
Laser Interferometer Space Antenna (LISA, see the LISA Consortium study
\citealt{Amaro-SeoaneEtAl2017} and also
\citealt{Amaro-SeoaneEtAl2013,Amaro-SeoaneEtAl2012}), or other space-borne
observatories alternatives, currently under study, Taiji \citep{taiji17} or
Tianqin \citep{tianqin16}. Moreover, IMRIs could be jointly detected with LISA and
ground-based detectors such as LIGO/Virgo, and LISA could warn them years in
advance and with seconds of precision \citep{Amaro-Seoane2018}.

Due to the fact that the IMBH is lighter than a supermassive black hole in a
galactic nucleus, we cannot assume anymore that it is fixed at the centre of
the system. This makes an analytical study of the problem arduous and
challenging.  In this article we model the potential formation of an IMRI in a
dense stellar system, a cluster, via the interactions between an IMBH and
stellar-mass black holes. The outcome of such process is varied and could lead
to different end states, including the formation of an IMRI, a stellar-mass
black hole binary, the ejection of one or both stellar-mass black holes, or
even a head-on collision between the compact objects.  To this end, we use a
direct-summation $N-$body code that includes post-Newtonian to account for the
periapsis shift and the loss of energy via gravitational radiation \citep[as
presented for the first time in][]{KupiEtAl06}, as well as the cluster
gravitational potential \textbf{(REF: Manuel, your first fork of ARCHAIN cited
here)}. We scan the parameter space, in particular the IMBH and stellar-mass
black hole masses, their orbital configuration, and the host cluster structural
properties with a set of 2,000 models.

In Section \ref{num} we present and summarise the numerical setup used to model
the interactions; in Section \ref{met} we discuss the main properties of the
simulations outputs, and the implications for GW astronomy; while Section
\ref{res} is devoted to the results and conclusions.
